### 2013-12-16 research/paper/figure/velocity_compare.tex
uploaded 12.01 11/11 2013
Committing changes and suggestions up to the 'effect of m*' section

    \put(0.03,0.95){ $\frac{V}{D}$} 	
    % \put(0.56,1.02){ $\frac{V}{D}$}

    \put(0.25,0.805){ $\frac{tU}{D}$} 	
    \put(0.73,0.805){ $\frac{tU}{D}$}

    \put(0.095,1.03){(a)}
    \put(0.565,1.03){(b)}

@@

  \caption{Time histories of velocity at two different $\zeta$ and $U^*$ which produce the same mean power ($1.2\times10^{-3}$). Data presented in (a) are at $\ustar=60$, $\zeta=0.075$ and (b) are at $\ustar=165$, $\zeta=0.175$. Both data sets were obtained using Quasi-steady state assumption using input $C_y$ parameters at $Re=165$. Shedding is evident in both signals as a high frequency fluctuation but the amplitude of the slower fluctuations remains constant in both cases.}
    \label{fig:time_hostory_velocity_same_power}
\end{figure}

>>>>

    \put(0.02,0.95){ $\displaystyle\frac{V}{D}$} 	
    % \put(0.56,1.02){ $\frac{V}{D}$}

    \put(0.25,0.80){ $\displaystyle\frac{tU}{D}$} 	
    \put(0.73,0.80){ $\displaystyle\frac{tU}{D}$}

    \put(0.095,1.035){(a)}
    \put(0.565,1.035){(b)}

@@

  \caption{Time histories of velocity at two different $\zeta$ and $U^*$ which produce the same mean power ($1.2\times10^{-3}$). Data presented in (a) are at $\ustar=60$, $\zeta=0.075$ and (b) are at $\ustar=165$, $\zeta=0.175$. Both data sets were obtained using Quasi-steady state assumption using input $C_y$ parameters at $Re=165$. Shedding is evident in both signals as a high frequency fluctuation but the amplitude of the slower fluctuations remains constant in both cases. \JL{KASUN: The different scale on the x-axes make these two plots difficult to compare. I think a more powerful comparison can be made by keeping the x-scale the same (include more periods of the case at $U^*=60$). Then it is clear that even though the oscillation happens over a longer time, the average power output is constant. You could also include a plot where you rescale time by the time scale I derived, $\tau=\tau=t(a_1/m^*)(U/D)$. I'm pretty certain that, apart from the high frequency shedding component, the two plots would lok the same using this time scale}}
    \label{fig:time_hostory_velocity_same_power}
\end{figure}

### 2013-12-16 research/paper/results/results-2.3.tex
Committing changes and suggestions up to the 'effect of m*' section

 Now the oscillator equation Eq.\eqref{final_equation_motion} is considered from a power perspective. The shedding forces can be neglected because the net effect is negligible as system oscillates at natural frequency which is far from shedding frequency, for the cases that exhibit galloping. It is obvious that the forcing term on LHS of the equation is only dependent on transverse velocity($\dot{y}$) which is essentially the input power of the system. On the RHS, the mechanical damping or system damping is the only term that takes out power at any instant. This could be expressed as the product of the damping force and the velocity ($P_d$). The inertia and the stiffness terms governs the frequency of the system but the forces associated with those terms are conservative forces (i.e there is zero net energy in or out of the system when averaged over a period). Therefore the system is governed by the transverse velocity rather than the natural frequency.

@@

%%%%%%%%%
The data presented in figure \ref{fig:uncollapsed_data} show that, regardless of the damping ratio $\zeta$, the trend of the response with $U^*$ is similar, suggesting the data can be collapsed by a suitable choice of parameters. These parameters can be found by considering the natural time scales of the equations of motion presented in equation \ref{final_equation_motion}, and by suitably non-dimensionalizing these equations.

@@

This timescale can then be used to non-dimensionalize the equation of motion, and to find the relevant dimensionless groups of the problem. If the non-dimensional time, $\tau$, is defined such that $\tau=t(a_1/m^*)(U/D)$, the equation of motion presented in equation \ref{final_equation_motion} can be non-dimensionalized as
\begin{equation}
  \label{eqn:eom_nondim}
  \ddot{Y} + \frac{m^{*2}}{a_1^2}\frac{k}{m}\frac{D^2}{U^2}Y = \left(\frac{1}{2} - \frac{m^*}{a_1}\frac{cD}{mU}\right)\dot{Y} \dots
\end{equation}
%%%%%%%%%

\JL{Stopped here}

 Using $U^*$ and $\zeta$ assumes that the system has a preferred frequency because the scale with the natural frequency of the system. The effect of fixing $\zeta$ and increasing $U^*$ actually decreases the damping constant for a fixed free-stream velocity. ($U^*=\frac{U}{f \times D}$, $\zeta= \frac{c}{2 m \omega_n}$ ). Both these effects leads to the multiple lines that are horizontally transpose when $\zeta$ is increased (Fig.\ref{fig:uncollapsed_data} (e) and (f)). Therefore the effect of $\zeta$ essentially scales up the damping coefficient for a fixed $U^*$.

 The data presented in Fig.\ref{fig:uncollapsed_data} for various damping ratios, $\zeta$, can be collapsed into a single line for a particular force characteristic curve (i.e $C_y$ vs $\theta$ curve). These collapsed curves were  obtained for the velocity amplitude  and power by plotting as functions of as a function of  the non dimensionalised  damping constant $\frac{c}{\rho\mathcal{A}U}$ 
(Fig \ref{fig:collpased_data} (a),(b),(c) and (d)).  This further emphasizes that the galloping system is not frequency dependent. It is possible to obtain a similar power output at different values of \ustar when the damping constant, $\frac{c}{\rho\mathcal{A}U}$, is kept fixed. An example of this case, as shown in Fig.\ref{fig:time_hostory_velocity_same_power}, clearly show that this is a result of similar velocity amplitudes between cases if one were to disregard the high frequencies due to shedding. As mentioned earlier, it is the transverse velocity that determines the energy provided by the fluid forcing and the mechanical damping.

@@

 Power could be expressed as the product of force and velocity. Therefore the transferred power form fluid-to-body could be expressed as $P_t=F_y\dot{y}$. Similarly the dissipated power due to the mechanical damping could be expressed as $P_d=(c\dot{y})\dot{y}$. The time average of these two quantities should be equal due to energy conservation, provided that the mechanical friction is neglected . Analysing the  time histories of $P_t $ and $P_d$ at key regions (Fig.\ref{fig:regions_1}) on the mean power vs $U^*$ provides a detailed explanation for the variation of the output power when the reduced velocity is increased. The key regions consists of region 1 where the $P_{mean}$ increases with \ustar, region 2 where $P_{mean}$ becomes maximum and region 3 where $P_{mean}$ decreases with \ustar. It has been established earlier that the damping factor is a function of $U^*$. Therefore it could be derived that $U^*$ is inversely proportional to damping coefficient. Hence the damping coefficient reduces when you move from region 1 to 3. Fig \ref{fig:lift_curves} (a) shows that $C_y$ and therefore instantaneous force rises until $4^\circ$ where it peaks and then falls and at around $6^\circ$ becomes negative. The maximum amount of power that could be transferred occurs near the peak region. At the region where the instantaneous force becomes negative it will be opposing the velocity $\dot{y}$. Data at $\zeta=0.1$, $m^*=40$ and \reynoldsnumber=165 (Fig.\ref{fig:power_time_histories}) are analysed as and example.

@@

 At region 1 where $U^*=90$ the damping constant is high and a clear sinusoidal signal could be observed for both $P_d$ and $P_t$ in Fig.\ref{fig:power_time_histories} (a). Fig.\ref{fig:power_time_histories} (d) and (g) shows that $\theta$ is in line or in phase with $F_y$.  The velocity amplitude in this case is small and the equivalent incident angle within the range where the hydrodynamic force increases with the incident angle (i.e. $0<\theta \leq 4^\circ$ in Fig.\ref{fig:lift_curves} (a)) Hence both $P_d$ and $P_t$ becomes sinusoidal. In this case, power output is limited by the low fluid forces present at low incident angles. In other words damping is significantly high and extracts a lot of power that the velocity amplitude could not grow where the forcing  is significant to produce high level of power.

  At region 3 ($U^*= 400$) `$c$' is low in comparison with region 1 and 2 which leads to a low mean power output. Fig.\ref{fig:power_time_histories} (c) shows that $P_t$ becomes negative over some portion of the cycle. This is caused by te high velocity amplitude leads to the equivalent incident angle $\theta$, in this case to exceed the range where $C_y$ is positive (i.e. $0<\theta<6^\circ$ in Fig.\ref{cy ploynomial} (a)). In this portion of the cycle the hydrodynamic force actually opposes the direction of travel and power is transferred from the structure to the fluid during those times. From and energy perspective, the mechanical damping is not sufficient to remover the energy transferred from the fluid to the structure during other times of the cycle because $\frac{c}{\rho\mathcal{A}U}$ is substantially low. Therefore this excess energy is transferred back to the fluid as depicted by the negative region of $P_d$ in Fig.\ref{fig:power_time_histories} (c)

At region 2 ($U^*=165$). $P_t$ is not a pure sinusoidal signal. However, the  signal remains periodic. From the time history graph of $P_t$, two `peaks' are present in a single half cycle (Fig \ref{fig:power_time_histories} (b)). In this case, the velocity amplitude actually exceeds the equivalent incident angle where the hydrodynamic forces peaks (i.e. $\theta=4^\circ$ in \ref{cy ploynomial} (a)). The dips in $P_d$ between the two peaks approximately correspond to the time where the transverse velocity is higher than 0.07 and $F_y$ is decreasing with increasing transverse velocity. The mean power output is at its maximum. This is due to the fact that this region is a best compromise between region 1 and 3. The damping is substantially high to obtain a high power output while not too high to allow the induced angle of attack to enter the region where the forcing opposes the direction of travel.

>>>>

 % Now the oscillator equation Eq.\eqref{final_equation_motion} is considered from a power perspective. The shedding forces can be neglected because the net effect is negligible as system oscillates at natural frequency which is far from shedding frequency, for the cases that exhibit galloping. It is obvious that the forcing term on LHS of the equation is only dependent on transverse velocity($\dot{y}$) which is essentially the input power of the system. On the RHS, the mechanical damping or system damping is the only term that takes out power at any instant. This could be expressed as the product of the damping force and the velocity ($P_d$). The inertia and the stiffness terms governs the frequency of the system but the forces associated with those terms are conservative forces (i.e there is zero net energy in or out of the system when averaged over a period). Therefore the system is governed by the transverse velocity rather than the natural frequency.

@@

The data presented in figure \ref{fig:uncollapsed_data} show that, regardless of the damping ratio $\zeta$, the trend of the response with $U^*$ is similar, suggesting the data can be collapsed by a suitable choice of parameters. These parameters can be found by considering the natural time scales of the equations of motion presented in equation \ref{final_equation_motion}, and by suitably non-dimensionalizing these equations.

@@

This timescale can then be used to non-dimensionalize the equation of motion, and to find the relevant dimensionless groups of the problem. If the non-dimensional time, $\tau$, is defined such that $\tau=t(a_1/m^*)(U/D)$, the equation of motion presented in equation \ref{final_equation_motion} can be non-dimensionalized as
\begin{equation}
  \label{eqn:eom_nondim}
  \ddot{Y} + \frac{m^{*2}}{a_1^2}\frac{kD^2}{mU^2}Y = \left(\frac{1}{2} - \frac{m^*}{a_1}\frac{cD}{mU}\right)\dot{Y} \dots,
\end{equation}

where the coefficients can be regrouped into combinations of non-dimensional groups, and rewritten as
\begin{equation}
  \label{eqn:eom_nondim_regroup}
  \ddot{Y} + \frac{k^{*2}m^{*2}}{a_1^2}Y = \left(\frac{1}{2} - \frac{c^*m^*}{a_1}\right)\dot{Y} \dots,
\end{equation}

where $k^*=kD^2/mU^2$ is a non-dimensional stiffness parameter, and $c^*=cD/mU$ is a non-dimensional damping parameter.

Equation \ref{eqn:eom_nondim_regroup} shows there are four non-dimensional parameters that play a role in setting the response of the system. These are the stiffness $k^*$, the damping $c^*$, the mass ratio $m^*$, and the rate of change in the aerodynamic force with respect to angle of attack when the body is at the equilibrium position, $a_1$. The grouping of these parameters into two groups in equation \ref{eqn:eom_nondim_regroup} which arise by non-dimensionalising using the natural time scale of the galloping system, suggests there are two groups that dictate the response: $\Gamma_1 = k^{*2}m^{*2}/a_1^2$ and $\Gamma_2 = c^*m^*/a_1$. For a given geometry and Reynolds number, $Gamma_1$ can be thought of as a combined mass-stiffness, whereas $\Gamma_2$ can be thought of a a combined mass-damping parameter. As it is assumed that during galloping the stiffnes plays only a minor role, $\Gamma_2$ seems a likely parameter to collapse the data presented in figure \ref{fig:uncollapsed_data}.

Figure \ref{fig:collapsed_data} presents the same data as shown in figure \ref{fig:uncollapsed_data}, but in terms of the parameter $\Gamma_2$. The collapse, especially of the velocity amplitude and power output data, is excellent. This result shows that it is possible to obtain a similar power output at different values of \ustar\ or $\zeta$ when the mass-damping constant, $\Gamma_2$, is kept fixed. An example of this is shown in Fig.\ref{fig:time_hostory_velocity_same_power}, where the time history of two cases with clearly different $U^*$ and $\zeta$ values, but the same value of $\Gamma_2=$\JL{KASUN: put the value of the mass-damping (your parameter $c/\rho\mathcal{A}U$) that you used here and in the figure caption}, are plotted. The figure clearly shows that the common power output is a result of similar velocity amplitudes between cases if one were to disregard the high frequencies due to shedding.

@@

Power can be expressed as the product of force and velocity. Therefore the transferred power fromm the fluid to the body can be expressed as $P_t=F_y\dot{y}$, as shown in equation \ref{power_alt}. Similarly the dissipated power due to the mechanical damping could be expressed as $P_d=(c\dot{y})\dot{y}$ as shown in equation \ref{power}. The time average of these two quantities should be equal due to energy conservation, provided that the mechanical friction losses are neglected. The mean power vs $U^*$ provides a detailed explanation for the variation of the output power when the reduced velocity is increased. The key regions consists of region 1 where the $P_{mean}$ increases with \ustar, region 2 where $P_{mean}$ becomes maximum and region 3 where $P_{mean}$ decreases with \ustar. Time histories of $P_t $ and $P_d$ at each of these key regions are presented in figure \ref{fig:power_time_histories}.

\JL{KASUN: why have you established that it is the damping, not $U^*$, that is important, but here you persist with presenting results in terms of $U^*$? How do these regions correspond to regions on the curve in terms of your damping parameter (that I have labelled $\Gamma_2$)? Can you define these three regions in terms of this new parameter?}

It has been established earlier that the damping factor is a function of $U^*$. Therefore it could be derived that $U^*$ is inversely proportional to damping coefficient. Hence the damping coefficient reduces when you move from region 1 to 3. Fig \ref{fig:lift_curves} (a) shows that $C_y$ and therefore instantaneous force rises until $4^\circ$ where it peaks and then falls and at around $6^\circ$ becomes negative. The maximum amount of power that could be transferred occurs near the peak region. At the region where the instantaneous force becomes negative it will be opposing the velocity $\dot{y}$. Data at $\zeta=0.1$, $m^*=40$ and \reynoldsnumber=165 (Fig.\ref{fig:power_time_histories}) are analysed as and example.

\JL{KASUN: I don't understand the point of the above paragraph. What are you trying to say? That maximum power output occurs when the tranverse velocity is such that the lift force is a maximum?}

@@

At region 1 where $U^*=90$ the damping constant is high and a clear sinusoidal signal is observed for both $P_d$ and $P_t$ in figure \ref{fig:power_time_histories}(a). Figures \ref{fig:power_time_histories}(d) and  \ref{fig:power_time_histories}(g) show that equivalent incident angle $\theta$ (which for small values, is proportional to the transverse velocity of the body) is in phase with $F_y$.  The velocity amplitude in this case is small and $\theta$ is within the range where the hydrodynamic force increases with the incident angle (i.e. $0<\theta \leq 4^\circ$ as shown in figure \ref{fig:lift_curves}(a)). According to equation \ref{power_alt}, these conditions are suitable for high power output. However in this case, the power output is limited by the high damping which limits the amplitude of the oscillation.

At region 3 ($U^*= 400$) the damping is low in comparison with region 1 and 2. While this may lead to larger oscillations, damping is required to dissipate power according to equation \ref{power}. Therefore, the low damping in this region leads to a low mean power output. Fig.\ref{fig:power_time_histories} (c) shows that $P_t$ becomes negative over some portion of the cycle. This is caused by the high velocity amplitude leading to the equivalent incident angle $\theta$ to exceed the range where $C_y$ is positive (i.e. $0<\theta<6^\circ$ as shown in figure\ref{cy ploynomial}(a)). In this portion of the cycle the hydrodynamic force actually opposes the direction of travel and power is transferred from the structure to the fluid during those times. From an energy perspective, the mechanical damping is not sufficient to removerthe energy transferred from the fluid to the structure during other times of the cycle because $\Gamma_2$ is substantially low. Therefore this excess energy is transferred back to the fluid as depicted by the negative region of $P_d$ in Fig.\ref{fig:power_time_histories}(c).

At region 2 ($U^*=165$), a balance is found between high and low values of damping. $P_t$ is not a pure sinusoidal signal, however the signal remains periodic. From the time history graph of $P_t$, two `peaks' are present in a single half cycle as shown if figure \ref{fig:power_time_histories}(b). In this case, the velocity amplitude actually exceeds the equivalent incident angle where the hydrodynamic forces peaks (i.e. $\theta=4^\circ$ in \ref{cy ploynomial} (a)). The dips in $P_d$ between the two peaks approximately correspond to the time where the transverse velocity is higher than 0.07 and $F_y$ is decreasing with increasing transverse velocity. The mean power output is at its maximum. This is due to the fact that this region is the best compromise between region 1 and 3. The damping is high enough to obtain a high power output while not too high to allow the induced angle of attack to enter the region where the forcing opposes the direction of travel.

\JL{KASUN: nothing to change here, just wanted to note that this paragraph is good. Clear, and to the point.}
